---
title: "London Research Landmarks"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

These examples were created for a Nature Toolbox article by Jeffrey Perkel, published 5 June 2018. [https://www.nature.com/articles/d41586-018-05331-6]

The following maps are created using the Leaflet R plugin. The Leaflet library produces interactive maps that allow users to pan, zoom, and click on points of interest.

I created and read in a table of latitude/longitude values in comma-separated values (CSV) format, using values extracted from Google Maps. Then I plot those points on a topographical map of London, and overlay a second data source from Macrostrat.org, which colors the map according to its geological features.

```{r echo=TRUE}
library (leaflet)

#
# In Google Maps, find a location, right click, select 'What's Here', and copy 
# the lat/long values. These values can be stored in a .csv file. I've added them 
# here to file for simplicity.
#

df <- read.csv(textConnection(
"Name,Lat,Long
Nature,51.533925,-0.121553
Francis Crick Institute,51.531877,-0.128767
University College London,51.524486,-0.133997
MRC Laboratory for Molecular Cell Biology,51.524435,-0.132495
King's College London,51.511573,-0.116083
Imperial College London,51.498780,-0.174888
Cambridge University,52.206960,0.115034
Oxford University,51.754843,-1.254302
Platform 9 3/4,51.532349,-0.123806
"))

# create a blank map
m <- leaflet() 

# mark points with blue circles, except Nature's offices in red...
m <- addCircleMarkers(m, lat=df$Lat, lng=df$Long, label=df$Name, 
               color = (ifelse (df$Name == "Nature", "red", "blue"))) 

# add an underlying basemap
m <- addProviderTiles(m, providers$OpenTopoMap) 

# center the view on London
m <- setView(m, -0.119,51.525, zoom = 8) 

# pull in MacroStrat tiles
m <- addTiles (m, 'https://tiles.macrostrat.org/carto/{z}/{x}/{y}.png', 
               options = tileOptions(opacity = 0.6))

# draw the map
m
```

Collect session information.
```{r}
sessionInfo()
```

